Holomorphic field theories and higher algebra

### 2023-08-17 highervertexops.tex
merge

\def\g{{\mathfrak g}}
\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\jou{{\mathtt{j}}}
\def\bgamma{{\mathbf{\gamma}}}
\def\bbeta{{\mathbf{\beta}}}
\def\CDO{{\cC\cD}}
\def\SU{{\rm SU}}
\def\sfa{\mathsf{a}}
\def\del{\partial}
\def\til{\Tilde}
\def\Sol{\text{Sol}}
\def\Spec{\text{Spec}}
\def\C{{\mathbb{C}}}
\def\RHom{{\rm RHom}}
\newcommand{\sff}{\mathsf{f}}
\newcommand{\sfe}{\mathsf{e}}
%\def\sfR{\mathsf{R}}
%\def\sfA{\mathsf{A}}

@@

Note that $p$ is not a map of algebras. 
For example,
\beqn
p(z_1 \lambda_1) = \frac12
\eeqn
as $z_1 \lambda_1$ admits a decomposition
\beqn
z_1 \lambda_1 = \frac12 (z_1 \lambda_1 - z_2 \lambda_2) + \frac12 .
\eeqn

@@

Define
\beqn
\op{Res} \colon \sfA \to \C[-1]
\eeqn
by the rules that $\op{Res}|_{\sfA^0} = 0$ and
\beqn
\op{Res} \left(f(z,\lambda) \omega \right) = p (f(z,\lambda)) (z=0) .
\eeqn
This residue extends to the completion to give a map
\beqn
\op{Res} \colon \Hat\sfA \to \C[-1] .
\eeqn

\begin{lem}
\label{lem:res} 
The following are true regarding the higher algebraic residue:
\begin{enumerate}
\item The residue is $\lie{sl}(2)$ invariant.
\item For $f \in \C[[z_1,z_2]]$ one has $\op{Res}(f \omega) = f(0)$.
\item For any $f \in \sfA$ one has $\op{Res}(\dbar f) = 0$.
Thus $\op{Res}$ is a cochain map.
\end{enumerate}
\end{lem}
\begin{proof}
Item (1) holds since $p$ is the $\lie{sl}(2)$-equivariant projection. 
For (2), note that for $f \in \C[z_1,z_2]$ that $p(f) = f$.

@@

\begin{prop}
The higher algebraic residue defines a cochain map
\beqn
\sfA \to \left(\sfA \, \d^2 z\right)^\vee [-1]
\eeqn 
which sends an element $a \in \sfA$ to the functional
\beqn
F_a \colon b \, \d^2 z \in \sfA\, \d^2 z \mapsto (-1)^{|a|} \Res(a b \, \d^2 z) .
\eeqn
\brian{weird sign, but I think it's there because of the $[-1]$ shift.}
When extended to the completed version of punctured affine space, this assignment defines an isomorphism of cochain complexes 
\beqn
F \colon \Hat\sfA \simeq \left(\Hat\sfA \, \d^2 z\right)^\vee [-1] .
\eeqn
\end{prop}
\begin{proof}
Suppose that $a \in \sfA^0$. 
Then
\begin{align*}
F_{\dbar a} (b \d^2 z) & = - \Res \left( (\dbar a) b \d^2 z \right) \\
& = - \Res\left( \dbar(a b \d^2 z) \right) + \Res \left(a \dbar b \d^2 z\right) \\
& = F_{a} (\dbar b \d^2 z) .
\end{align*}
In the final line we have used (3) from lemma \ref{lem:res}.
If $a \in \sfA^1$ then $F_{\dbar a}(b \d^2 z) = 0 = F_{a} (\dbar b \d^2 z)$ for all $b$.

@@

>>>>

\def\g{{\mathfrak g}}
\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\jou{{\mathtt{j}}}
\def\bgamma{{\mathbf{\gamma}}}
\def\bbeta{{\mathbf{\beta}}}
\def\CDO{{\cC\cD}}
\def\SU{{\rm SU}}
\def\sfa{\mathsf{a}}
\def\del{\partial}
\def\til{\Tilde}
\def\Sol{\text{Sol}}
\def\Spec{\text{Spec}}
\def\C{{\mathbb{C}}}
\def\RHom{{\rm RHom}}
\newcommand{\sff}{\mathsf{f}}
\newcommand{\sfe}{\mathsf{e}}
%\def\sfR{\mathsf{R}}
%\def\sfA{\mathsf{A}}
\newcommand{\laurent}{\Hat \sfA}
\newcommand{\sfb}{\mathsf{b}}
\newcommand{\sfc}{\mathsf{c}}

@@

Note that $p$ is not a map of algebras. 
For example,
\beqn
p(z_1 \lambda_1) = \frac12
\eeqn
as $z_1 \lambda_1$ admits a $\lie{sl}(2)$-decomposition
\beqn
z_1 \lambda_1 = \frac12 (z_1 \lambda_1 - z_2 \lambda_2) + \frac12 .
\eeqn

@@

Define
\beqn
\op{Res} \colon \sfA \to \C[-1]
\eeqn
by the rules that $\op{Res}|_{\sfA^0} = 0$ and
\beqn
\op{Res} \left(f(z,\lambda) \omega \right) = p (f(z,\lambda)) (z=0) .
\eeqn
This residue extends to the completion to give a map
\beqn
\op{Res} \colon \laurent \to \C[-1] .
\eeqn

\begin{lem}
\label{lem:res} 
The following are true regarding the higher algebraic residue:
\begin{enumerate}
\item The residue is $\lie{sl}(2)$ invariant.
\item For $f \in \C[[z_1,z_2]]$ one has $\op{Res}(f \omega) = f(0)$.
\item For any $f \in \laurent$ one has $\op{Res}(\dbar f) = 0$.
Thus $\op{Res}$ is a cochain map.
\end{enumerate}
\end{lem}
\begin{proof}
Item (1) holds since $p$ is the $\lie{sl}(2)$-equivariant projection. 
For (2), note that for $f \in \C[z_1,z_2]$ that $p(f) = f$.

@@

\begin{prop}\label{prop:res}
The higher algebraic residue defines a cochain map
\beqn
\sfA \to \left(\sfA \, \d^2 z\right)^\vee [-1]
\eeqn 
which sends an element $a \in \sfA$ to the functional
\beqn
F_a \colon b \, \d^2 z \in \sfA\, \d^2 z \mapsto (-1)^{|a|} \Res(a b \, \d^2 z) .
\eeqn
\brian{weird sign, but I think it's there because of the $[-1]$ shift.}
When extended to the completed version of punctured affine space, this assignment defines an isomorphism of cochain complexes 
\beqn
F \colon \Hat\sfA \simeq \left(\Hat\sfA \, \d^2 z\right)^\vee [-1] .
\eeqn
\end{prop}
\begin{proof}
Suppose that $a \in \sfA^0$. 
Then
\begin{align*}
F_{\dbar a} (b \d^2 z) & = - \Res \left( (\dbar a) b \d^2 z \right) \\
& = - \Res\left( \dbar(a b \d^2 z) \right) + \Res \left(a \dbar b \d^2 z\right) \\
& = F_{a} (\dbar b \d^2 z) .
\end{align*}
In the final line we have used (3) from lemma \ref{lem:res}.
If $a \in \sfA^1$ then $F_{\dbar a}(b \d^2 z) = 0 = F_{a} (\dbar b \d^2 z)$ for all $b$.

@@

\section{Algebraic free field realization}

Consider the vector space $V = \laurent$.
Via the residue pairing we have an identification $V^\vee = \laurent \, \d^2 z [1]$, see proposition \ref{prop:res}.
From hereon we use the notation $\laurent^{(2)} = \laurent \, \d^2 z$.
We will denote linear elements in the corresponding Poisson algebra
\beqn
\cP = \Sym \left(\laurent  \oplus \laurent^{(2)}[1] \right) ,
\eeqn
by $(\sfb , \sfc)$.
On these elements, the Poisson bracket reads
\beqn
\{\sfb ,\sfc\} = \Res(\sfb \sfc) .
\eeqn
The Lie algebra $\cP_2$ is spanned by symbols of the form $\sfb \sfb',\sfb \sfc, \sfc \sfc'$ where $\sfb,\sfb'\in \laurent$ and $\sfc,\sfc' \in \laurent^{(2)}[1]$.

The algebra $\cP$ is the free commutative algebra generated by $\laurent \oplus \laurent^{(2)}[1]$.
Alternatively, we can view, via the residue pairing, a linear element $(\sfb, \sfc) \in \laurent \oplus \laurent^{(2)}[1]$ as the functional
\beqn
(\sfb, \sfc) \colon \laurent^{(2)} [1] \oplus \laurent \to \C
\eeqn
defined by
\beqn
(\beta, \gamma) \mapsto (-1)^{|\sfb|} \Res(\sfb \beta) + (-1)^{|\sfc|} \Res(\sfc \gamma) .
\eeqn
Likewise, we can view a quadratic element in $\cP_2$ as a functional
\beqn
\Sym^2 \left(\laurent^{(2)} [1] \oplus \laurent \right) \to \C .
\eeqn
For example, the quadratic element $\sfb \sfc \in \cP_2$ corresponds to the quadratic functional
\beqn
\beta \gamma \mapsto (-1)^{|\sfb|+ |\sfc|} \Res (\sfb \beta) \Res(\sfc \gamma) ,
\eeqn
where $\beta \in \laurent^{(2)} [1]$ and $\gamma \in \laurent$.

Let $J$ be the map
\beqn
J \colon \sfA \to \cP_2
\eeqn
defined by sending an element $a \in \sfA$ to the quadratic functional
\beqn
J_a (\beta \gamma) = \Res(\beta a \gamma) .
\eeqn

\begin{lem}
The map $J$ is a cochain map.
Furthermore, for any $a,a' \in \sfA$ we have $\{J_a, J_{a'}\} = 0$.
Thus, if we endow $\sfA$ with the structure of an abelian dg Lie algebra then $J$ is a map of dg Lie algebras.
\end{lem}
\begin{proof}
For the first claim observe that
\beqn
J_{\dbar a} (\beta \gamma) = \Res(\beta (\dbar a) \gamma) = \pm \Res((\dbar \beta) a \gamma) \pm \Res(\beta a \dbar \gamma) = J_a (\dbar \beta \gamma) + (-1)^{|\beta|} J_a(\beta \dbar \gamma) 
\eeqn
by (3) in lemma \ref{lem:res}.
\end{proof}
